---
title: "Castration extends lifespans of both male and female by 15%"
author: "Michael Garratt, Jean-François Lemaître, Malgrozata Lagisz & Shinichi Nakagawa"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
    
    rmdformats::robobook:
      code_folding: hide
      code_download: true
editor_options: 
  chunk_output_type: console
# output:
#   html_document:
#     code_download: true
#     code_folding: hide
#     depth: 4
#     number_sections: no
#     theme:  cosmo # “default”, “cerulean”, “journal”, “flatly”, “darkly”, “readable”, “spacelab”, “united”, “cosmo”, “lumen”, “paper”, “sandstone”, “simplex”, “yeti”
#     toc: yes
  #   toc_float: yes
  #   toc_depth: 4
  # pdf_document:
  #   toc: yes
subtitle: Supplementary Material
#bibliography: references.bib
#biblio-style: "apalike"
#csl: ecol_lett.csl # we can change this
#link-citations: yes
#always_allow_html: true
---

```{r setup, include = FALSE}
#kniter seetting
knitr::opts_chunk$set(
message = FALSE,
warning = FALSE, # no warnings
cache = TRUE,# Cacheing to save time when kniting
tidy = TRUE
#fig.width = 9
)

# clearning up
rm(list=ls())
```

# Setting-ups

## Loading packages

```{r}
# packages ####

#ochaRd
# 
# install.packages("devtools")
# install.packages("tidyverse")
# #install.packages("metafor")
# install.packages("patchwork")
# install.packages("R.rsp")
# 
# devtools::install_github("daniel1noble/orchaRd", force = TRUE, build_vignettes = TRUE)
# remotes::install_github("rvlenth/emmeans", dependencies = TRUE, build_opts = "") 
# 
# #emmeans
# remotes::install_github("rvlenth/emmeans", dependencies = TRUE, build_opts = "")
# # metafor
# install.packages("remotes")
# remotes::install_github("wviechtb/metafor")

# loading
pacman::p_load(tidyverse,
               metafor,
               pander,
               stringr,
               ape,
               kableExtra,
               patchwork,
               here,
               lme4,
               readxl,
               metaAidR,
               rotl,
               orchaRd,
               emmeans,
               clubSandwich
)
```

## Custon functions

```{r}
# custom functions 

# function for getting lnRR for proportional data

lnrrp <- function(m1, m2, n1, n2) {
  # arcsine transforamtion
  asin_trans <- function(p) { asin(sqrt(p)) }
  # SD for arcsine distribution (see Wiki - https://en.wikipedia.org/wiki/Arcsine_distribution)
  var1 <- 1/8
  var2 <- 1/8
  # lnRR - with 2nd order correction
  lnrr <- log(asin_trans(m1)/asin_trans(m2)) + 
    0.5 * ((var1 / (n1 * asin_trans(m1)^2)) - (var2 / (n2 * asin_trans(m2)^2)))	
  
  var <- var1 / (n1 * asin_trans(m1)^2) + var1^2 / (2 * n1^2 * asin_trans(m1)^4)  + 
    var2 / (n2 * asin_trans(m2)^2) + var2^2 / (2 * n2^2 * asin_trans(m2)^4) 
  
  invisible(data.frame(yi = lnrr , vi = var))
}

# function for getting lnRR for mean data

lnrrm <- function(m1, m2, n1, n2, sd1, sd2) {
  # lnRR - with 2nd order correction
  lnrr <- log(m1/m2) + 
    0.5 * (((sd1^2) / (n1 * m1^2)) - ((sd2^2) / (n2 * m2^2)))	
  
  var <- sd1^2 / (n1 * m1^2) + sd1^4 / (2 * n1^2 * m2^4)  + 
    sd2^2 / (n2 * m2^2) + sd2^4 / (2 * n2^2 * m2^4) 
  
  invisible(data.frame(yi = lnrr , vi = var))
}

```

# Datasets 

- I will put two data sets here to show with meta-data

## Main dataset and meta-data

```{r, eval =FALSE}
# # getting the data and formating some variables (turning chraracter vectors to factors)
# read_csv(here("data/lit_search.csv"), na = "NA") %>%
#    mutate_if(is.character, as.factor) %>%  kable("html") %>%
#   kable_styling("striped", position = "left")
```

## Sub-dataset and meta-data

```{r, eval = FALSE}
# # getting the data and formating some variables (turning chraracter vectors to factors)
# full_data <- read_csv(here("data/2021-09-01-source-data-dat.csv"), na = "NA") %>% 
#    mutate_if(is.character, as.factor)
# 
# # dataset to compare the same cophylogenies between the two methods
# 
# full_pair <- read_csv(here("data/2020-08-12-paried.csv"), na = "NA") %>% 
#    mutate_if(is.character, as.factor)
# 
# # making a scrollable table
# kable(full_data, "html") %>%
#   kable_styling("striped", position = "left") %>%
#   scroll_box(width = "100%", height = "500px")
```

A. __authors__: The authors of the study and the date (citation form).

B. __year__: The year of publication of the study.

C. __host_tax_broad__:	Separation of the host group according to broader taxonomic units (e.g. vertebrate, invertebrate, microbe, plant).

D. __symbiont_tax_broad__:	Separation of the symbiont group according to broader taxonomic units (e.g. vertebrate, invertebrate, microbe, plant).

E. __symbiont_euk__*:	Whether the symbiont is eukaryotic (state ='yes'), or prokaryotic (state='no').


```{r}

```

## Data processing

### Main data
```{r}
# loading data ####
dat_full1 <- read_csv(here("data", "dat_15082021.csv"), na = c("", "NA"))


# deleting unusable rows #####
# excluding Vasectomy 
dat_full %>% filter(is.na(Treatment_lifespan_variable) == FALSE) %>% filter(Type_of_sterilization != "Vasectomy") -> dat

im(dat)
dim(dat_full)
# separating two kinds

effect_type <- str_detect(dat$Lifespan_parameter, "Me")


# effect-level ID

dat$Effect_ID <- 1:nrow(dat)
dat$Phylogeny <- sub(" ", "_",  dat$Species_Latin)
dat$Effect_type <- effect_type


# we create a longer data format

dat1 <- dat
dat2 <- dat
```

### Sub-data


