---
title: "Castration extends lifespans of both male and female by 15%"
author: "Michael Garratt, Jean-François Lemaître, Malgrozata Lagisz & Shinichi Nakagawa"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
    
    rmdformats::robobook:
      code_folding: hide
      code_download: true
editor_options: 
  chunk_output_type: console
# output:
#   html_document:
#     code_download: true
#     code_folding: hide
#     depth: 4
#     number_sections: no
#     theme:  cosmo # “default”, “cerulean”, “journal”, “flatly”, “darkly”, “readable”, “spacelab”, “united”, “cosmo”, “lumen”, “paper”, “sandstone”, “simplex”, “yeti”
#     toc: yes
  #   toc_float: yes
  #   toc_depth: 4
  # pdf_document:
  #   toc: yes
subtitle: Supplementary Material
#bibliography: references.bib
#biblio-style: "apalike"
#csl: ecol_lett.csl # we can change this
#link-citations: yes
#always_allow_html: true
---

ML NOTES:  
 - I kept top chunks and added new ones for summarising the dataset AFTER some pre-processing (by SN)  
 - Data from "dat_19092021.csv" file  
 - custom functions could be moved to an external source R file  

```{r setup, include = FALSE}
#kniter seetting
knitr::opts_chunk$set(
message = FALSE,
warning = FALSE, # no warnings
cache = TRUE,# Cacheing to save time when kniting
tidy = TRUE
#fig.width = 9
)

# clearning up
rm(list=ls())
```

# Setting-up

## Loading packages

```{r load packages}
# packages ####

#ochaRd
# 
# install.packages("devtools")
# install.packages("tidyverse")
# #install.packages("metafor")
# install.packages("patchwork")
# install.packages("R.rsp")
# 
# devtools::install_github("daniel1noble/orchaRd", force = TRUE, build_vignettes = TRUE)
# remotes::install_github("rvlenth/emmeans", dependencies = TRUE, build_opts = "") 
# 
# #emmeans
# remotes::install_github("rvlenth/emmeans", dependencies = TRUE, build_opts = "")
# # metafor
# install.packages("remotes")
# remotes::install_github("wviechtb/metafor")

# loading
pacman::p_load(tidyverse,
               metafor,
               pander,
               stringr,
               ape,
               kableExtra,
               patchwork,
               here,
               lme4,
               readxl,
               metaAidR,
               rotl,
               orchaRd,
               emmeans, 
               naniar,
               GoodmanKruskal,
               ggplot2,
               ggalluvial,
               clubSandwich
)
```

## Custom functions

```{r custom functions}

# function for getting lnRR for proportional data

lnrrp <- function(m1, m2, n1, n2) {
  # arcsine transforamtion
  asin_trans <- function(p) { asin(sqrt(p)) }
  # SD for arcsine distribution (see Wiki - https://en.wikipedia.org/wiki/Arcsine_distribution)
  var1 <- 1/8
  var2 <- 1/8
  # lnRR - with 2nd order correction
  lnrr <- log(asin_trans(m1)/asin_trans(m2)) + 
    0.5 * ((var1 / (n1 * asin_trans(m1)^2)) - (var2 / (n2 * asin_trans(m2)^2)))	
  
  var <- var1 / (n1 * asin_trans(m1)^2) + var1^2 / (2 * n1^2 * asin_trans(m1)^4)  + 
    var2 / (n2 * asin_trans(m2)^2) + var2^2 / (2 * n2^2 * asin_trans(m2)^4) 
  
  invisible(data.frame(yi = lnrr , vi = var))
}

# function for getting lnRR for mean data

lnrrm <- function(m1, m2, n1, n2, sd1, sd2) {
  # lnRR - with 2nd order correction
  lnrr <- log(m1/m2) + 
    0.5 * (((sd1^2) / (n1 * m1^2)) - ((sd2^2) / (n2 * m2^2)))	
  
  var <- sd1^2 / (n1 * m1^2) + sd1^4 / (2 * n1^2 * m2^4)  + 
    sd2^2 / (n2 * m2^2) + sd2^4 / (2 * n2^2 * m2^4) 
  
  invisible(data.frame(yi = lnrr , vi = var))
}

```

### Datasets   

[Shinichi will put two data sets here to show with meta-data?]   

(Main dataset and meta-data)   
```{r, eval = FALSE}
# # getting the data and formating some variables (turning chraracter vectors to factors)
# read_csv(here("data/lit_search.csv"), na = "NA") %>%
#    mutate_if(is.character, as.factor) %>%  kable("html") %>%
#   kable_styling("striped", position = "left")
```

(Sub-dataset and meta-data)  

```{r, eval = FALSE}
# # getting the data and formating some variables (turning chraracter vectors to factors)
# full_data <- read_csv(here("data/2021-09-01-source-data-dat.csv"), na = "NA") %>% 
#    mutate_if(is.character, as.factor)
# 
# # dataset to compare the same cophylogenies between the two methods
# 
# full_pair <- read_csv(here("data/2020-08-12-paried.csv"), na = "NA") %>% 
#    mutate_if(is.character, as.factor)
# 
# # making a scrollable table
# kable(full_data, "html") %>%
#   kable_styling("striped", position = "left") %>%
#   scroll_box(width = "100%", height = "500px")
```


### Data processing - for main data  

```{r load main data}
# loading data ####
dat_full1 <- read_csv(here("data", "data_19092021.csv"), na = c("", "NA"))
names(dat_full1)
#dim(dat_full1) #118 rows

# deleting unusable rows #####
# excluding Vasectomy 
# TODO this is important!!!!!
dat_full1 %>% filter(is.na(Treatment_lifespan_variable) == FALSE) %>% 
  filter(Type_of_sterilization != "Vasectomy") %>% 
  mutate_if(is.character, as.factor) -> dat
#dim(dat) #116 (i.e. removed 2 rows)

# separating two kinds of effect size (longevity vs. mortality measures)
effect_type <- ifelse(str_detect(dat$Lifespan_parameter, "Me"), "longevity", "mortality")
#table(effect_type)

#add new variables
dat$Effect_ID <- 1:nrow(dat) # effect-level ID
dat$Phylogeny <- sub(" ", "_",  dat$Species_Latin) #Pjylogeny species names (for the tree) based on Latin names
dat$Effect_type <- as.factor(effect_type)

#extract publication year from Study variable
dat$Year <- as.numeric(gsub(".*?([0-9]+).*", "\\1", dat$Study)) #equivalent to parse_number(dat$Study)but does not create problems for NA  
```

## Summarise the dataset  

Note that ES have not been calculated yet (use prelim.R code to do so).  

Look at the overall pattern of missing data.  

```{r missing data, echo = FALSE}
#vis_miss(dat) #mostly data is missing at comment fields

vis_miss(dat) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 3), 
        plot.margin = margin(t = 0.5, r = 2, b = 1, l = 1, unit = "cm")) +
  ggtitle("Missing data overall") #no mising values
```


### General summary   

 - Number of effect sizes: `r length(unique(dat$Effect_ID))`   
 - Number of studies: `r length(unique(dat$Study))`   
 - Publication years: from `r min(dat$Year, na.rm = TRUE)` to `r max(dat$Year, na.rm = TRUE)`    
 - Simple list of studies as short references: `r unique(dat$Study)`     
 - Number of species: `r length(unique(dat$Species_Latin))`     
 - Simple list of species Latin names (also stored in "Phylogeny" variable): `r unique(dat$Species_Latin)`    
 - Number of data points for females: `r length(which(dat$Sex=="Female"))`, and data points form males: `r length(which(dat$Sex=="Male"))`    
 

```{r key variables simple tables, eval = FALSE}
#Key variables: Sex, Gonads_removed, Wild_or_semi_wild, Controlled_treatments, Effect_type, Maturity_at_treatment_ordinal (ordinal with NA), 
#names(dat)

#check how crossed with Sex (relevant to both sexes or not?)
table(dat$Species_Latin, dat$Sex)
table(dat$Wild_or_semi_wild, dat$Sex)
table(dat$Gonads_removed, dat$Sex) #always Yes in Males
table(dat$Controlled_treatments, dat$Sex)
table(dat$Maturity_at_treatment_ordinal, dat$Sex) #table(dat$Maturity_at_treatment)
table(dat$Effect_type, dat$Sex)

#check how crossed with Wild_or_semi_wild
table(dat$Gonads_removed, dat$Wild_or_semi_wild)
table(dat$Controlled_treatments, dat$Wild_or_semi_wild)
table(dat$Maturity_at_treatment_ordinal, dat$Wild_or_semi_wild) 
table(dat$Effect_type, dat$Wild_or_semi_wild)

#check how crossed with Gonads_removed
table(dat$Wild_or_semi_wild, dat$Gonads_removed)
table(dat$Controlled_treatments, dat$Gonads_removed)
table(dat$Maturity_at_treatment_ordinal, dat$Gonads_removed) 
table(dat$Effect_type, dat$Gonads_removed)

#check how crossed with Controlled_treatments
table(dat$Wild_or_semi_wild, dat$Controlled_treatments)
table(dat$Gonads_removed, dat$Controlled_treatments)
table(dat$Maturity_at_treatment_ordinal, dat$Controlled_treatments)
table(dat$Effect_type, dat$Controlled_treatments)

#check how crossed with Effect_type
table(dat$Wild_or_semi_wild, dat$Effect_type)
table(dat$Gonads_removed, dat$Effect_type)
table(dat$Controlled_treatments, dat$Effect_type)
table(dat$Maturity_at_treatment_ordinal, dat$Effect_type)
```

### Key predictors (moderators)   

Visualise missing data in 6 key variables (moderators):   
 - Sex,   
 - Wild_or_semi_wild,   
 - Maturity_at_treatment_ordinal,  
 - Gonads_removed,   
 - Controlled_treatments,   
 - Effect_type  
  
```{r key variables missing values, echo = FALSE}
#create a data subset wth key variables only
dat_key <- dat %>% select(Sex, Wild_or_semi_wild, Maturity_at_treatment_ordinal, Gonads_removed, Controlled_treatments, Effect_type)

vis_miss(dat_key) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 3), 
        plot.margin = margin(t = 0.5, r = 2, b = 1, l = 1, unit = "cm")) +
  ggtitle("Missing data in key variables") #32% mising values in Maturity_at_treatment_ordinal only
```

Visualise pairwise associations between 6 key variables:   

```{r key variables associations, echo = FALSE}
#names(dat_key)
#useGoodman and Kruskal’s τ measure of association between categorical predictor variables (function from package GoodmanKruskal: https://cran.r-project.org/web/packages/GoodmanKruskal/vignettes/GoodmanKruskal.html)
GKmatrix <- GKtauDataframe(subset(dat_key))
#par(mar=c(1,1,2,4))
plot(GKmatrix)
```


Alluvial diagrams for key predictor variables (moderators) - two groups of 3 moderators, grouped by similarity:  

 A. For Sex, Gonads_removed, Maturity_at_treatment_ordinal (changed NA to 0)  


```{r alluvial plots A, results = 'hide'}
#use ggalluvial (https://cran.r-project.org/web/packages/ggalluvial/vignettes/ggalluvial.html)

# create a frequency table for first 3 moderator variables
# freq_1 <- as.data.frame(table(dat$Sex, dat$Gonads_removed, dat$Maturity_at_treatment_ordinal)) %>% rename(Sex = Var1, Gonads_removed = Var2, Maturity_at_treatment_ordinal = Var3)
# is_alluvia_form(as.data.frame(freq_1), axes = 1:3, silent = TRUE)
# #freq_1 %>% filter(Freq != 0) %>% arrange(desc(Freq)) #collapesd table of values, without 0s

 
# ggplot(data = freq_1,
#   aes(axis1 = Sex, axis2 = Gonads_removed, axis3 = Maturity_at_treatment_ordinal, y = Freq)) +
#   geom_alluvium(aes(fill = Sex)) +
#   geom_stratum(aes(fill = Sex))+
#   geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
# #theme_minimal() +
#   theme_void() +
#   theme(legend.position = "none",
#         plot.title = element_text(hjust = 0.5, vjust = 3),
#         axis.title.x = element_text(),
#         axis.text.x = element_text(face="bold")) +
#   scale_x_discrete(limits = c("Sex", "Gonads removed", "Maturity class"), expand = c(0.15, 0.05), position = "top") +
#   scale_fill_brewer(palette = "Set3") +
#   ggtitle("A. Subjects sex, manipulation type and maturity class")

#as above but with, but with Gonads_removed as first column, and different colours
# ggplot(data = freq_1,
#   aes(axis1 = Gonads_removed, axis2 = Sex, axis3 = Maturity_at_treatment_ordinal, y = Freq)) +
#   geom_alluvium(aes(fill = Gonads_removed)) +
#   geom_stratum(aes(fill = Gonads_removed))+
#   geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
#   theme_minimal() +
# #  theme_void() +
#   theme(legend.position = "none",
#         plot.title = element_text(hjust = 0.5, vjust = 3),
#         axis.title.x = element_text(),
#         axis.text.x = element_text(face="bold")) +
#   scale_x_discrete(limits = c("Gonads_removed", "Sex", "Maturity class"), expand = c(0.15, 0.05), position = "top") +
#   scale_fill_brewer(palette = "Pastel2") +
#   ggtitle("A. Subjects sex, manipulation and maturity ")

#NOTE: all rows with NA in Maturity class are removed from the plot 
#recode NA as 0
dat$Maturity_at_treatment_ordinal2 <- dat$Maturity_at_treatment_ordinal
dat$Maturity_at_treatment_ordinal2[is.na(dat$Maturity_at_treatment_ordinal2)] <- 0
# create a frequency table for first 3 moderator variables
freq_1 <- as.data.frame(table(dat$Sex, dat$Gonads_removed, dat$Maturity_at_treatment_ordinal2)) %>% rename(Sex = Var1, Gonads_removed = Var2, Maturity_at_treatment_ordinal = Var3)
is_alluvia_form(as.data.frame(freq_1), axes = 1:3, silent = TRUE)
#freq_1 %>% filter(Freq != 0) %>% arrange(desc(Freq)) #collapesd table of values, without 0s

 
# ggplot(data = freq_1,
#   aes(axis1 = Sex, axis2 = Gonads_removed, axis3 = Maturity_at_treatment_ordinal, y = Freq)) +
#   geom_alluvium(aes(fill = Sex)) +
#   geom_stratum(aes(fill = Sex))+
#   geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
# #theme_minimal() +
#   theme_void() +
#   theme(legend.position = "none",
#         plot.title = element_text(hjust = 0.5, vjust = 3),
#         axis.title.x = element_text(),
#         axis.text.x = element_text(face="bold")) +
#   scale_x_discrete(limits = c("Sex", "Gonads removed", "Maturity class"), expand = c(0.15, 0.05), position = "top") +
#   scale_fill_brewer(palette = "Set3") +
#   ggtitle("A. Subjects sex, manipulation type and maturity class")

#as above but with, but with Gonads_removed as first column, and different colours
ggplot(data = freq_1,
  aes(axis1 = Gonads_removed, axis2 = Sex, axis3 = Maturity_at_treatment_ordinal, y = Freq)) +
  geom_alluvium(aes(fill = Gonads_removed)) +
  geom_stratum(aes(fill = Gonads_removed))+
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
#  theme_void() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5, vjust = 3),
        axis.title.x = element_text(),
        axis.text.x = element_text(face="bold")) +
  scale_x_discrete(limits = c("Gonads_removed", "Sex", "Maturity class"), expand = c(0.15, 0.05), position = "top") +
  scale_fill_brewer(palette = "Pastel2") +
  ggtitle("A. Subjects sex, manipulation and maturity ")
```

 B. For Wild_or_semi_wild, Controlled_treatments, Effect_type (Outcome type):   

```{r alluvial plots B, results = 'hide'}
# create a frequency table for next 3 moderator variables
freq_2 <- as.data.frame(table(dat$Wild_or_semi_wild, dat$Controlled_treatments, dat$Effect_type)) %>% rename(Wild_or_semi_wild = Var1, Controlled_treatments = Var2, Effect_type = Var3)
is_alluvia_form(as.data.frame(freq_2), axes = 1:3, silent = TRUE)
#freq_2 %>% filter(Freq != 0) %>% arrange(desc(Freq)) #collapesd table of values, without 0s

ggplot(data = freq_2,
  aes(axis1 = Wild_or_semi_wild, axis2 = Controlled_treatments, axis3 = Effect_type, y = Freq)) +
  geom_alluvium(aes(fill = Wild_or_semi_wild)) +
  geom_stratum(aes(fill = Wild_or_semi_wild))+
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
#  theme_void() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5, vjust = 3),
        axis.title.x = element_text(),
        axis.text.x = element_text(face="bold")) +
  scale_x_discrete(limits = c("Wild_or_semi_wild", "Controlled_treatments", "Outcome type"), expand = c(0.15, 0.05), position = "top") +
  scale_fill_brewer(palette = "Pastel1") +
  ggtitle("B. Experimental settings and outcome type")

#names(dat_key)
```


C. For Gonads_removed, Wild_or_semi_wild, Controlled_treatments, Effect_type (Outcome type):   

```{r alluvial plots C, results = 'hide'}
# create a frequency table for next 3 moderator variables
freq_3 <- as.data.frame(table(dat$Gonads_removed, dat$Wild_or_semi_wild, dat$Effect_type)) %>% rename(Gonads_removed = Var1, Wild_or_semi_wild = Var2, Effect_type = Var3)
is_alluvia_form(as.data.frame(freq_3), axes = 1:3, silent = TRUE)
#freq_2 %>% filter(Freq != 0) %>% arrange(desc(Freq)) #collapesd table of values, without 0s

ggplot(data = freq_3,
  aes(axis1 = Gonads_removed, axis2 = Wild_or_semi_wild, axis3 = Effect_type, y = Freq)) +
  geom_alluvium(aes(fill = Gonads_removed)) +
  geom_stratum(aes(fill = Gonads_removed))+
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  theme_minimal() +
#  theme_void() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5, vjust = 3),
        axis.title.x = element_text(),
        axis.text.x = element_text(face="bold")) +
  scale_x_discrete(limits = c("Gonads_removed", "Wild_or_semi_wild", "Outcome type"), expand = c(0.15, 0.05), position = "top") +
  scale_fill_brewer(palette = "Set3") +
  ggtitle("C. Manipulation, subject and outcome types")

#names(dat_key)
```

This just shows a different combination of predictor variables (moderators) from plots A and B.      